Place of death in NRS COVID-19 deaths: data processing for analysis

### 2021-05-18 covid_nrs_place_of_death_data.Rmd
added a table of excess deaths

>>>>

# Excess deaths

## Excess deaths by place of death

```{r}
total_deaths_overall %>%
  mutate(excess_deaths = deaths_all_causes - annual_deaths_min)

excess_deaths_2020_table <-
  past_deaths_overall %>%
  group_by(year, place_of_death) %>%
  summarise(annual_deaths = sum(number_of_deaths), .groups = "drop") %>%
  filter(year==2019) %>%
  pivot_wider(names_from = year, names_prefix = "annual_deaths_", values_from = annual_deaths) %>%
  left_join(
    total_deaths_overall %>% filter(year==2020), by = "place_of_death"
    ) %>%
  mutate(
    excess_deaths_vs_2019 = deaths_all_causes - annual_deaths_2019,
    # excess_deaths_vs_2019_prop = excess_deaths_vs_2019 / sum(excess_deaths_vs_2019),
    excess_deaths_vs_mean = deaths_all_causes - annual_deaths_mean,
    excess_deaths_vs_min = deaths_all_causes - annual_deaths_min,
    excess_deaths_vs_max = deaths_all_causes - annual_deaths_max
  ) %>%
  select(year, place_of_death, annual_deaths_2019, annual_deaths_mean, deaths_all_causes, matches("excess")) %>%
  mutate(disaggregated_place_of_death = place_of_death != "All") %>%
  group_by(disaggregated_place_of_death) %>%
  mutate(
    proportion_excess_deaths_vs_2019 = excess_deaths_vs_2019 / sum(excess_deaths_vs_2019),
    proportion_excess_deaths_vs_mean = excess_deaths_vs_mean / sum(excess_deaths_vs_mean)
  ) %>%
  ungroup %>%
  select(-disaggregated_place_of_death)

excess_deaths_table_2019 <-
  excess_deaths_2020_table %>%
  select(place_of_death, annual_deaths_2019, deaths_all_causes, excess_deaths_vs_2019, proportion_excess_deaths_vs_2019) %>%
  mutate(
    proportion_excess_deaths_vs_2019 = scales::percent(proportion_excess_deaths_vs_2019, accuracy = 0.1)
    ) %>%
  mutate_if(is.numeric, ~format(.x, big.mark=",")) %>%
  rename(
    "Place of death" = place_of_death,
    "2019 deaths" = annual_deaths_2019,
    "2020 deaths (all causes)" = deaths_all_causes,
    "Excess deaths" = excess_deaths_vs_2019,
    "Proportion of excess deaths by place" = proportion_excess_deaths_vs_2019
  )

write.xlsx(x = list("Excess w.r.t. 2019" = excess_deaths_table_2019), file = paste0(dir_outputs,"/excess_deaths_Scotland_compared_to_2019.xlsx"))
```
